### 2014-06-23 tex/author-title-etc.tex
minor edits

%!TEX root = flock-comment-main.tex

%%%%% THIS IS THE SECTION WHERE THE AUTHOR PUTS IN ALL OF THEIR TITLE AND AFFILIATION
%%%%% INFORMATION AND  A FEW OTHER THINGS
\newcommand{\myTitle}{A statistically-grounded interpretation of {\sc flock}}

% redefine this to make author list. Note affiliation symbols are done manually.
% All caps tends to look better
\newcommand{\myAuthors}{ ERIC C.~ANDERSON$^{*,\dagger,\S}$ and PATRICK D. BARRY$^\ddag$}

% redefine to make the affiliation list.  Note symbols are done manually
\newcommand{\myAffiliations}{$^*$Fisheries Ecology Division, 
    Southwest Fisheries Science Center, National Marine Fisheries Service, NOAA,
    110 Shaffer Road,
    Santa Cruz, CA 95060, USA, $^\dagger$Department of Applied Math and Statistics, University of
California, Santa Cruz, CA, $^\ddag$University of Alaska, Fairbanks, School of Fisheries and Ocean Science,
17101 Point Lena Loop Rd.,
Juneau, AK 99801, USA }

% the email address for the corresponding author
\newcommand{\myEmailAddress}{eric.anderson@noaa.gov}
\newcommand{\myEmailFootnote}{$^\S$}

% here you can put your very own copyright notice
\newcommand{\myCopyright}{\copyright US Federal Government work in the public domain in the USA}

% here you can put a running title (a short title that goes on the left of the 
% even pages)
\newcommand{\myRunningTitle}{{\sc flock} as a special case of {\sc structure}}

% and here you put the running author (short listing of authors for the
% upper right header on the odd pages)
\newcommand{\myRunningAuthor}{Anderson \& Barry}

%%%% DONE WITH AUTHOR/TITLE/ETC INFORMATION DEFINITION