### 2014-08-21 tex/abstract.tex
Fixed some mathematical sections then submitted.
On a last reading I found some bone-headed errors in
the mathematical section --- I left the \ell subscript
on things when it shouldn't have been.  So, I cleaned
that up and fixed it.

Then included referees in the cover letter and fixed a
few typos and submitted it.

This is the version that was submitted.

%!TEX root = flock-comment-main.tex

      We show that the the algorithm in the program {\sc flock} \citep{Duc&Tur2009} can be
interpreted as an estimation procedure using  
a model that is essentially identical to the {\sc structure} 
\citep{Pritchardetal2000} model with no admixture and non-correlated 
allele frequency priors. The computational routine in {\sc flock} is 
equivalent to searching for the maximum-{\em a-posteriori}
estimate of this restricted {\sc structure} model via a simulated 
annealing algorithm with an extreme cooling 
schedule (namely, the exponent on the objective function~$\rightarrow \infty$).  By analysis of a large, real data set we
confirm that {\sc flock} yields results similar to {\sc structure} with the no-admxixture
model and uncorrelated allele frequency prior with more variability 
from run to run. Interpreting {\sc flock}'s algorithm as a special case of the model in 
{\sc structure} should aid in  understanding the program's output and behavior.