### 2015-02-11 tex/main-body-text.tex
edited conclusion

%!TEX root = flock-comment-main.tex

\section*{Introduction}
Methods for the unsupervised clustering of genotypes have received 
considerable attention in the molecular ecology literature.  
The best known example of this class of methods is {\sc structure} 
\citep{Pritchardetal2000}, with over 12,000 citations identified on 
Google Scholar.  A variety of other 
clustering methods have been developed that are all closely related to 
{\sc structure}, for example {\sc NewHybrids} \citep{And&Tho2002}, {\sc 
BayesAss+} \citep{Wil&Ran2003}, and {\sc baps} 
\citep{Coranderetal2004}. An overview of the similarities between several of these methods can be 
found in \citet{Anderson2009PGAC}.

The recently introduced software {\sc flock} \citep{Duc&Tur2009} has
been described
as a ``non-Bayesian method [that] 
differs substantially from previous 
clustering algorithms'' \citep[][p.~1333]{Duc&Tur2009}. As reported
in a later paper, ``{\sc flock} is very different from  other clustering 
programs. It does not sample the space of partitions through small 
random step walks as in MCMC, and it does not try to optimize some 
target function, such as HWLE\@. Briefly stated, it is not based on a 
probabilistic search algorithm. On the contrary, {\sc flock} is 
entirely deterministic'' \citep[][p.~736]{Duc&Tur2012}.

However, here we identify 
the close relationship between {\sc flock}'s algorithm and the program
{\sc structure}.  Although the algorithm in {\sc flock}
is, as the authors point out, deterministic (apart from a random initial allocation of 
individuals to clusters), it can be interpreted as a 
limiting case of a probabilistic search algorithm applied to one of {\sc structure}'s
models.  More specifically, the {\sc flock} 
algorithm is a special case of the simulated annealing
algorithm for finding the Bayesian maximum-{\em a-posteriori}
estimate from a marginalized form of the {\sc structure} model with no
admixture and non-correlated allele frequencies.

Being able to interpret {\sc flock} in this fashion should help users
interpret output from {\sc flock} and predict when it might and might not
give substantially different results than {\sc structure}. Below we describe
the correspondence between the programs in more detail, and then, for illustration, compare
the results obtained from {\sc flock} and {\sc structure} on a large, real data
set.

\section*{Comparison of methods}
We start with a succinct mathematical description of the {\sc structure}
model, then describe the {\sc flock} algorithm, and finally explain
the close relationship between the two.

\subsection*{{\sc structure}}
In the {\sc structure} model with no admixture (which, it should be pointed out, 
is not the default model in {\sc structure}), the unknown ``subpopulation'' that the 
$i\thh$ individual ($i=1,\ldots,N)$ belongs to is denoted by $Z_i \in \{1,\ldots,K\}$, 
where $K$ is the 
number of subpopulations (or clusters, as they are often referred to).  
In a diploid individual from cluster $k$, the 
allelic types of the two gene copies at the $\ell\thh$ locus are assumed
to be drawn independently from the vector of allele frequencies at
locus $\ell$ in subpopulation $k$,  $\theta_{k\ell}=(\theta_{k\ell 1},\ldots,\theta_{k\ell A_\ell})$, where 
$A_\ell$ 
denotes the number of alleles observed in the data set at locus $\ell$.
Hence, if $Y_{i\ell}$ denotes a vector of length $A_\ell$ whose components denote the 
number
of copies of each of the $A_\ell$ alleles at locus $\ell$ in a diploid individual $i$, and 
$Z_i=k$, then $Y_{i\ell}$ follows the multinomial distribution of two trials with
$A_\ell$  components and cell probabilities given by the allele frequencies: 
\begin{equation}
(Y_{i\ell}~|~Z_i=k) \sim \mathrm{Mult}_{A_\ell}(2, \theta_{k\ell}).
\end{equation}
In the {\sc structure} model without physical linkage, the genotypes at the loci are assumed to
be independent of one another, so the probability of the genotype data at
all $L$ loci---$Y_i=(Y_{i1},\ldots,Y_{iL})$---is simply a product of multinomial 
probabilities.
In the uncorrelated allele frequencies model, the prior on each $\theta_{k\ell}$ is a
Dirichlet distribution with parameters $(\lambda_{k\ell1},\ldots,\lambda_{k\ell A_
\ell})$,
which are usually all set to a value such as 1 or $1/A_\ell$.

After initializing the unknown allele frequencies, $\theta = (\theta_1,\ldots,\theta_K)$, to randomly drawn values,
$\theta^{(0)}$, inference in the model proceeds by sampling from the joint posterior of the 
$Z_i$'s and $\theta$ using Gibbs sampling.  That is,
at iteration $t = 0, 1, 2, \ldots$:
\begin{enumerate}
\item each $Z^{(t)}_i$ is updated to $Z^{(t+1)}_i$ by sampling a value from 
the full conditional distribution of $Z_i$ given
$Y_i$ and $\theta^{(t)}$ (the current estimate of the allele frequencies).  This
distribution is found using Bayes' theorem. In {\sc structure}'s formulation, the 
prior probability that $Z_i=k$ is $1/K$ for all $k=1,\ldots, K$.     
\item $\theta^{(t)}$ is updated to $\theta^{(t+1)}$ from its full conditional distribution.  For each 
cluster
$k$ and locus $\ell$, the full conditional for $\theta_{k\ell}$ is independently a 
Dirichlet distribution with parameters $\lambda_{k\ell j} + \#(k,\ell,j)$, for 
$j=1,\ldots, A_\ell$,
where $\#(k,\ell,j)$ is the number of alleles of type $j$ at locus $\ell$ observed in 
individuals with $Z^{(t+1)}_i = k$.   
\end{enumerate}

It will be useful for our comparison with {\sc flock} to point out that another way 
of pursuing inference in this model would be to first integrate out the 
Dirichlet priors on the allele frequencies and then sample from the posterior
for the $Z_i$'s by Gibbs sampling.  When $\theta$ is integrated out, the genotypes 
of the individuals are no longer conditionally independent (they were originally
independent {\em conditional} on $\theta$ and the $Z_i$'s), so the calculation of 
the full conditional distribution of $Z_i$ becomes somewhat more involved,
but can be computed by the following reasoning. First, the 
conditional distribution of $Y_{i\ell}$, given that individual $i$ is from subpopulation $k$
(\ie $Z_i=k$), depends on the 
cluster memberships and genotypes at locus $\ell$ of all the remaining individuals,
which we denote by  $Z_{(-i)}$ and $Y_{(-i)\ell}$, respectively.
This full conditional for $Y_{i\ell}$ follows 
a Compound Dirichlet Multinomial distribution (CDM):
\begin{eqnarray}
\lefteqn{(Y_{i\ell}~|~Z_i=k,~Z_{(-i)},~Y_{(-i)\ell}) \sim} \label{eq:cdm}\\
& & \mathrm{CDM}(\lambda_{k\ell 1} + \#_{(-i)}(k,\ell,1), \ldots,
\lambda_{k\ell A_\ell} + \#_{(-i)}(k,\ell,A_\ell)), \nonumber
\end{eqnarray}
where $\#_{(-i)}(k,\ell,j)$ is the number of alleles of type $j$ at locus $\ell$
found in individuals {\em other than individual $i$} that currently
belong to cluster $k$. Thus, calculating the product of (\ref{eq:cdm}) across all the loci
($\ell = 1,\ldots,L$) gives the full conditional multilocus genotype probability 
for individual $i$:
\begin{eqnarray}
\lefteqn{P(Y_{i}~|~Z_i=k,~Z_{(-i)},~Y_{(-i)}) =}  \label{eq:cdm-multi-prob} \\
& &  \prod_{\ell=1}^L P(Y_{i\ell}~|~Z_i=k,~Z_{(-i)},~Y_{(-i)\ell}). \nonumber
\end{eqnarray}
Computing (\ref{eq:cdm-multi-prob}) for 
for each value of $Z_i=k \in 
\{1,\ldots,K\}$ 
and normalizing to sum to one gives the full conditional distribution for 
$Z_i$:
\begin{equation}
P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)})~~,~~k=1,\ldots,K.
\label{eq:fc}
\end{equation}
A new value of $Z_i$ would be drawn from (\ref{eq:fc}) if doing Gibbs sampling in this
version of the {\sc structure} model in which $\theta$ has been integrated out.  In 
fact,
this is the approach (with a slightly different prior weight on the $Z_i$'s) taken to 
update 
$Z_i$ in both {\sc hwler} \citep{Pel&Mas2006} and {\sc structurama} \citep{Hue&And2007} 
when not
proposing changes to the number of clusters.

\subsection*{{\sc flock}}
Here we translate the procedure given by \citet{Duc&Tur2009} for the 
{\sc flock} algorithm into a specification in terms of the variables
defined in the previous section.  We provide the description for a case in which
it is assumed there are $K$ clusters.

The first step of the flock algorithm is initialization, during which 
the program randomly allocates each of the $N$ individuals to one of 
$K$ clusters.  Then a number of reallocation steps are performed.  During every one of 
these steps, each individual is given the chance to be reallocated 
(\ie moved to a different cluster).  This is done on the basis of 
maximum likelihood: as \citeauthor{Duc&Tur2009} say (2009, p.~1335), ``Re-allocations are 
performed following multilocus maximum likelihood (Paetkau et al. 1995).''

From that description, it is not clear how \citeauthor{Duc&Tur2009} treat 
alleles that appear in the focal
individual but do not appear within any other individuals within a cluster.
The original approach of \citet{Paetkauetal1995} merely added 0.01 or some other
small value to each allele frequency that was 0, and then renormalized 
the allele frequencies to sum to 1.  In later works, Paetkau and colleagues
also employed the ``Bayesian'' approach of \citet{Ran&Mou1997} in their 
software {\sc geneclass2}  \citep{Piryetal2004}.  If \citeauthor{Duc&Tur2009}
use the approach of \citet{Ran&Mou1997}, then the $i\thh$ individual will
be reallocated to whichever cluster gives the highest value to 
$P(Y_{i}~|~Z_i=k,~Z_{(-i)},~Y_{(-i)})$, which is 
exactly the probability
defined in (\ref{eq:cdm-multi-prob}).  
If \citeauthor{Duc&Tur2009} use the simpler ``add 0.01 and renormalize''
formulation of \citet{Paetkauetal1995}, then their reallocations are based
on maximizing a likelihood that, although it may not be formally identical to
the probability defined in (\ref{eq:cdm-multi-prob}), is nearly identical to it.

It is worth pointing out that, since the
{\sc structure} model without admixture assumes a uniform prior over 
the $K$ different clusters, $P(Y_{i}~|~Z_i=k,~Z_{(-i)},~Y_{(-i)})$
is exactly proportional to $P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)})$, which 
we have shown is the
full conditional distribution for $Z_i$ in the {\sc structure} model after
integrating out $\theta$.  Thus, we have shown that, doing Gibbs
sampling for $Z_i$ in  the {\sc structure} model in which $\theta$ has been
integrated out would involve sampling a new value of $Z_i$ from the 
probability distribution,
\[
P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)})
\]
and that the updates that
{\sc flock} makes to each $Z_i$ involve assigning to $Z_i$
\begin{equation}
\arg\max_k P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)}),
\end{equation}
\ie the value of $k$ that maximizes $P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)})$
(or, as stated above, a nearly identical function of $k$, depending on how
the program {\sc flock} treats alleles that are not observed in certain clusters).

\subsection*{The {\sc flock} algorithm and {\sc structure}}
The goal of {\sc structure}'s MCMC algorithm is to sample from the space of 
$Z$'s in proportion to their posterior probability.  If, all that was desired
was a point estimate of the $Z_i$'s that maximized the posterior probability,
then one standard approach to finding that maximum-{\em a-posteriori}, or MAP, 
estimate, would be to use simulated annealing \citep{Kirkpatricketal1983}.
To do simulated annealing in the {\sc structure} model with no admixture and
no correlated allele frequency prior, after integrating out the unknown
allele frequencies, the updates for each $Z_i$ would be made from the 
distribution
\[
\biggl[P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)})\biggr]^\beta
\]
which is the same full conditional distribution as the marginalized {\sc structure}
model (equation~\ref{eq:fc}), raised
to the power $\beta$.  In typical applications of simulated annealing,
$\beta$ is set to a starting value less than 1, and, as the algorithm proceeds,
$\beta$ is increased according to a ``cooling schedule''\citep{Hajek1988}, until
the algorithm finds a local maximum in the posterior probability surface.  
When $\beta$ is small, the process can easily traverse ``valleys''
in the posterior probability surface and has a better chance of searching
more of the space for the maximum value.  As $\beta$ gets larger, it becomes
more probable that the process will
move uphill on the probability surface, tending toward the local
maximum near where it currently is.

It should now be clear that the algorithm in {\sc flock} is a simulated annealing 
algorithm for the {\sc structure} model in which the cooling schedule is
``start with $\beta\rightarrow\infty$ and leave it there for the duration of 
the algorithm.''  As such, we can expect that it may be more susceptible 
to becoming trapped in local modes of the posterior probability. 
In the following section we show that {\sc structure} and {\sc flock} find similar
solutions in the analysis of a large microsatellite data set, though, as expected,
{\sc flock} seems to find the same solution less consistently than does {\sc structure}.

\section*{Comparison of {\sc structure} and {\sc flock}}
{\sc flock}, initially designed to map admixture in hybrid populations 
\citep{Duc&Tur2009}, was recently proposed as an alternative to 
{\sc structure} for estimating the number of populations from a collection with 
unsampled source populations \citep{Duc&Tur2012}. In doing so, 
{\sc flock} relies on individuals with like genotypes clustering together 
through an iterative process. It stands to reason then, a reasonable way 
to gauge how well the algorithm accomplishes the main objective 
of clustering individuals into their correct groups is to simulate data of known 
population structure with varying levels of differentiation and 
evaluate how well it does assuming the correct number of populations. 
The implementation of {\sc flock} as a VBA procedure (macro) in 
Microsoft Excel precludes a rigorous evaluation of its behavior due 
to the inability to efficiently batch process. Consequenty, we adapted 
the source code from {\sc structure} as describe above to create {\sc flockture}
(code provided on GitHub repository https://github.com/eriqande/flockture) 
in order to accurately assess the strengths and weaknesses of the {\sc flock} 
algorithm. First, using simulated data we demonstrated that {\sc flock} 
and {\sc flockture} are essentially equivalent programs.
Second, we simulated and analyzed datasets with {\sc flockture} 
and three models in {\sc structure} (the admixture and correlated 
allele frequency (A-C) model, the no-admixture and correlated 
allele frequency (NA-C) model, and the no-admixture and non-correlated 
allele frequency (NA-NC) model) to help better understand
under what conditions, if any, running {\sc flock} may be preferable to {\sc structure} .

For each set of runs, {\sc flock} reports the 
normalized likelihood values (the exponentiated log-likelihood value for an individual's 
membership to each of $K$ clusters, scaled to sum to one) for a run belonging to the 
longest plateau - refered to as the 'best run' by \citep{Duc&Tur2012} (see below).
These values can be compared to the $q_i$ values
returned by {\sc structure}, which, in the no-admixture case, are posterior 
probabilities of cluster membership. \citet{Duc&Tur2012} describe a method 
for using {\sc flock} to estimate, $K$, the number of
subpopulations. The user must specify both the number of runs 
(different initial allocations of individuals 
to each cluster) as well as the number of iterations. (In each iteration every individual 
has the opportunity to be reallocated 
to a different cluster.) At the conclusion of each run, the log-likelihood value for 
membership to all $K$ clusters is computed
and the log likelihood difference (LLOD), the difference between the 
cluster with the highest likelihood 
and second highest likelihood, is calculated for each individual. {\sc flock} 
identifies runs that have converged to the same solution by detecting 
that they have the same mean LLOD score, and, in 
{\sc flock} parlance, a series of runs that have converged to the same 
solution are termed a ``plateau.'' The plateau record is a list of 
how many runs converged to the same allocation
of individuals. \citet{Duc&Tur2012} developed two stopping 
rules that they use to estimate $K$. \citet{Duc&Tur2012} advocate 
running  {\sc flock} until either a single plateau of
$\geq 6$ runs is reached or four consecutive $K$ values yield no 
plateaus. If the first stopping condition is satisfied, the $K$ for which 
there is a single plateau of $\geq 6$ runs is the point estimate for $K$. 
Alternatively, if multiple plateaus are observed with at least one plateau 
of $\geq 6$ runs, the largest value of $K$ that satisfies 
this conditon should be taken as the lower bound estimate of $K$. 
If no plateaus of $\geq 6$ runs are observed, $K$
is declared {\em undecided}.

To compare {\sc flock} and {\sc flockture} 4 datasets composed of 5 populations
were simulated using a coalescent based program (code provided on GitHub repository 
https://github.com/eriqande/flock-comment/simdata). Each of the 5 populations
were composed of 250, 250, 400, 550, and 550 individuals with 15 loci genotypes
and an average pairwise $F_{ST}$ of 0.03, 0.02, 0.01, and 0.005 among
populations. Both programs were run nine times at $K=5$ with the default setting for all 
run parameters (initial partition = random, number of iterations = 20, number of runs = 50, 
LLOD threshold = 0). We judged {\sc flockture} as an accurate 
representation of {\sc flock} based on the
following criteria: (1) Clustering of individuals into corrected population as
 defined by a zero-one loss function, (2) individual $q_i$ values, and (3) plateau records. 
All runs were performed on a JNCS D685 with an i7 3.40GHz processor and 8GB of 
RAM running 64-bit Windows 7.

Once we were confident that no major differences in inference existed between 
{\sc flock} and {\sc flockture} we used {\sc flockture} to analyze 48 datasets for two marker types 
(microsatellites and single nucleotide polymorphisms (SNPs))  that encompassed a broader 
range of population differentiation. The 96 datasets were simulated as above (five populations 
composed of 250, 250, 400, 550, and 550 individuals with 15 loci genotypes for the microsatellites
and 96 loci for the SNPs)
with average pairwise $F_{ST}$ values ranging between 0.004 and 0.045. 
 {\sc flockture} was run with the default run parameters and each run of 
{\sc structure} consisted of a 5,000 sweeep burn-in followed by a 20,000 sweep sample
from the posterior distribution. All runs were performed on a Macbook with a 2.3 GHz i7 processor
and 16GB of RAM running OS X ver. 10.7.5. The program {\sc clumpp} \citep{Jak&Ros2007} 
was used to relabel clusters among different runs,
while the program {\sc district} \citep{Rosenberg2004} was used to plot individual \textit{$q_i$} values.

\subsection*{{\sc flock} and {\sc flockture}} 
{\sc flock} and {\sc flockture} provided similar clustering of individuals among all four 
simulated datasets. The zero-one loss across all datasets had eqivalent results (Fig.~\ref{fig:FvFloss}).
When genetic differentiation is high both programs correctly apportion the majority of individuals and 
as genetic differentiation degrades among populations both programs show corresponding declines in 
assignment ability. Neither the mean nor range of loss between {\sc flock} and {\sc flockture} differs
substantially among the simulated datasets. These assignments were dependent on the normalized likelihood
values ($q_i$), which while they were comparable may be slightly biased toward the bounds (0,1)
for {\sc flock} (Fig.~\ref{fig:FvFqi}). In an attempt to understand the 
deviations being observed we took a subset of the 
first dataset (29 individuals from population 1 and 21 individuals from population 2) and initiated 
{\sc flock} with all individuals coming from population 1 starting in reference population 1 and all 
individuals from population 2 starting in reference population 2. {\sc flock} was run a second time
with the same starting clusters, but with the reference population labels switched. While the final allocation
of individuals matched the mean LLOD was slightly different (5.045009 vs. 5.028084). As a result, comparisons
between {\sc flock} and {\sc flockture} were limited to random initial starting points and should not 
be identical, but should have displayed equivalent behavior.

Plateaus were only observed for the dataset 
with the highest pairwise $F_{ST}$ value $(0.03)$. Plateau records were similar, although 
the length of each plateau record was slightly longer for {\sc flockture} and on average had more runs converging on
the largest plateau (Supp. Table 1). Multiple runs of {\sc flockture} demonstrated that the differences are well within the
variability that we would expect from randomizing the initial grouping of individuals to 
initiate the algorithms.

\begin{figure*}
\centering
\begin{minipage}{.45\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{images/Figures-Pat/FlockvFlocktureLoss.pdf}
  \captionof{figure}{Zero-one loss function for {\sc flock} and {\sc flockture} for 9 replicate runs of simulated datasets comprised 
of five populations with varying levels of population differentiation:  $F_{ST} = $ (a) 0.03, (b) 0.02, (c) 0.01, and (d) 0.005.}
  \label{fig:FvFloss}
\end{minipage}%
\hfill
\begin{minipage}{.45\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{images/Figures-Pat/FlockvFlockture_qi.pdf}
  \captionof{figure}{Normalized likelihood values for {\sc flock} and {\sc flockture} for a single run of simulated
data with average pairwise $F_{ST} = $ 0.03. }
  \label{fig:FvFqi}
\end{minipage}
\end{figure*}

Run times for both programs were relatively short, albeit in the case of {\sc flock} 
non-trivial. {\sc flock} averaged 1966.67 ($\pm 64.32$) seconds
over all runs, while {\sc flockture} averaged 9.83 ($\pm 0.21$) seconds. Run times showed a 
general increase as population structure became more subtle, however {\sc flock} did not 
demonstrate a linear increase. % this is surprising to me as it is coin the same thing irrespective of FST
% What I assume is happening is that as FST increases you have a lot more individuals being shuffled around
% at later iterations leading to increased process times. Seems reasonable.

\subsection*{{\sc flockture} and {\sc structure}} 
As expected {\sc flockture} and {\sc structure} both perform extremely well on datasets with
well differentiated populations. Infrequently, at higher levels of population differentiation the 
NA-NC and NA-C models in {\sc structure} exhibited large losses (Fig.~\ref{fig:uSatLoss})
resulting from the assignment of individuals from two groups into a single cluster while 
individuals from another group were separated into two clusters
with intermediate $q_i$ values (Fig.~\ref{fig:FvSdistruct}  A). This phenomena 
(only present in a single microsatellite dataset) disappears as population differentiation 
declines. As the subtlety of genetic structure 
increased loss for both programs increases and differences in the performance of
each model with different marker types emerged.
For the microsatellite dataset we observed that both programs
produce roughly equivalent results at higher levels of population differentiation. At intermediate 
values, the variance of the loss incurred by each model increases and certain models do marginally
better in assignment. No model did consistently better than others, until $F_{ST}$ drops 
below 0.0136 after which the loss for {\sc flockture} was markedly larger as was the among run
variance. The drastic increase in variance among runs was also observed with the SNP datasets. 
 {\sc flockture} did, however, have consistenlty smaller loss than  {\sc structure} at high to intermediate 
values of $F_{ST}$. Once $F_{ST}$ fell below 0.016 {\sc structure} began correctly assigning a 
higher proportion of individuals.

The number and length of plateaus {\sc flockture} encountered quickly 
diminished as the amount of differentiation among populations decreased. For the microsatellite datasets
only one run at $F_{ST} = 0.0339$ produced a single plateau of length greater than six. The other eight 
replicate runs yielded plateau records varying in length from two to eight all of which had at 
least one plateau larger than thirteen. When $F_{ST}$ was larger than 0.0337 no plateau record
contained at least one plateau of length six or greater. No plateaus were observed for any run when 
$F_{ST}$ was larger than 0.02187. Results from the SNP datasets
were concordant. No run of {\sc flockture} resulted in a single plateau of length greater than six.
Only simulated datasets with pairwise $F_{ST}$ values greater 0.0323 had plateau records with a 
plateau greater than six. Despite the lack of plateaus observed for many of the simulated datasets 
graphs of the normalized likelihood values from the {\sc flockture} run with the highest 
likelihood were suggestive of population structure.

Plots of the normalized likelihood values from {\sc flockture} can be useful in discerning the presence of 
structure.  In situations where only a single run of {\sc flockture} produced a plateau of length 2 
(Fig.~\ref{fig:FvSdistruct} B) the variation in assignment of individuals among runs was indistinguisable
and results were highly concordant with those of {\sc structure}. As population differentiation degrades 
and plateaus are not observed, the run with the highest likelihood from {\sc flockture} still yields
similar clusters as {\sc structure} but with more variation among runs (Fig.~\ref{fig:FvSdistruct} C \& D).
As pairwise $F_{ST}$ decreases and the variation among {\sc flockture} runs increases the normalized likelihood
values for assignment remain high in contrast to the $q_i$ values in each of the three models of {\sc structure}.

  \begin{figure*}
\centering
\includegraphics[width=.9\linewidth]{images/Figures-Pat/uSatloss.pdf}%Need to make a greyscale version
  \caption{Zero-one loss for nine runs of {\sc flockture} and the admixture and correlated allele 
  frequency (A.C) model, the no-admixture and correlated allele frequency (NA.C) model, 
and the no-admixture and non-correlated allele frequency (NA.NC) model in {\sc structure}. Forty-eight 
datasets were simulated with varying levels of population differentation. A coalescent based program
was used to generate 15 loci microsatellite genotypes for 2000 individuals comprising five populations.}
  \label{fig:uSatLoss}
\end{figure*}

 \begin{figure*}
\centering
  \includegraphics[width=.9\linewidth]{images/Figures-Pat/SNPloss.pdf}%Need to make a greyscale version
  \caption{Zero-one loss for nine runs of {\sc flockture} and the admixture and correlated allele 
  frequency (A.C) model, the no-admixture and correlated allele frequency (NA.C) model, 
and the no-admixture and non-correlated allele frequency (NA.NC) model in {\sc structure}. Forty-eight 
datasets were simulated with varying levels of population differentation. A coalescent based program
was used to generate 96 loci single nucleotide polymorphism (SNP) genotypes for 2000 
individuals comprising five populations.}
  \label{fig:SNPloss}
\end{figure*}

 \begin{figure*}
\centering
  \includegraphics[width=.8\linewidth]{images/Figures-Pat/FlockturevStructureDistruct.pdf}
  \caption{Plots of $q_i$ values from {\sc structure} and the normalized likelihood values from
  {\sc flockture} made with the program {\sc distruct}. Each horizontal plot represents one of nine 
 runs of the programs  at $K$ = 5. Three models were run in structure: the admixture and correlated allele 
  frequency (A-C) model, 
the no-admixture and correlated allele frequency (NA-C) model, 
and the no-admixture and non-correlated allele frequency (NA-NC) model. Each simulation is comprised of
 five populations of sizes 250, 250, 400, 550, and 550 individuals with average pairwise $F_{ST}$ values
 indicated above each plot.}
  \label{fig:FvSdistruct}
\end{figure*}

\section*{Conclusions}
In the past decade, the use of genetic markers to identify population structure and to attempt inference of
the number of genetic clusters ($K$) in a collection has dramatically increased in the fields of ecology, evolution, 
epidemiology, and conservation
genetics. The most commonly used program for doing this is {\sc structure} \citep{Pritchardetal2000,Falushetal2003}, 
a full-featured program with many options that is widely-used, well-tested, and universally regarded as a standard tool amongst molecular ecologists. 
Estimating $K$ is a difficult task, and because it is often based on individual assignment we 
evaluated these programs on their ability to correctly assign indivdiuals at the true value
of $K$. We have shown here that the model incorporated by {\sc flock} is essentially equivalent to the no-admixture model 
with non-correlated allele frequencies within the program {\sc structure}. 
We adapted the source code of {\sc structure} to create {\sc flockture} and show that
 {\sc flock} and {\sc flockture} are essentially equivalent programs. 
In comparing results from {\sc flockture} and {\sc structure} 
on simulated datasets with varying levels of differnetiation among populations
we find that the two programs give similar results.

Though we observed similar results
between the two programs at large values of $F_{ST}$, 
{\sc flockture}'s `best-run' solutions consistently assigned fewer individuals correctly
and became somewhat inconsistent at lower values of $F_{ST}$.
This is likely a consequence of the low population divergence and also
the large size of the search space over possible partitions. 
The likelihood surface defined on partitions of this space will often
be multimodal---possessing numerous local peaks and troughs.
The peak that  {\sc flockture}'s algorithm finds is entirely determined by the initial, random, allocation of 
individuals, as the algorithm proceeds deterministically after the initial random allocation.
By having a large number of individuals
that can be arranged into 5 groups defines a large partion space.
Decreasing $F_{ST}$ leads to more homogenous 
populations whose multilocus genotypes have a smaller difference in 
likelihood of assigning to any given cluster. This could act by increasing the number of
local peaks, leading {\sc flockture} settle on many different solutions. The problem of getting 
caught in local modes is not unique to {\sc flock}---{\sc structure} also can converge to 
different solutions in complex spaces; however our results suggest this is a bigger problem for {\sc flockture}.

One of the advantages of {\sc flock} over other methods for genetic clustering
is reported to be the short processing time for each run of {\sc flock}.
While \citet{Duc&Tur2012} observed faster processing times (relative to other methods), 
their comparision was based on setting {\sc flock} to perform 50 runs and 20 iterations for each 
$K$ starting at 2 and ending when one of the stopping conditions was met, and computing times for that analysis 
were compared to ten iterations of the admixture and correlated allele frequency model in {\sc structure} using 
a 50,000 sweep burn-in followed by a 200,000 sweep sample 
from the posterior for values of $K$ from 1 to three more than the true 
value of $K$. It is difficult to reliably assess differences in program run times when there is no clear consensus on what 
constitutes a reasonable comparison with respect to run length and number of runs, however, we show that if 
{\sc flock} were to be implemented in a coding language rather than a VBA routine in microsoft excel
it would be approximately 200 times faster.

While there has been lengthy 
discourse over the choice of estimators for $K$ in {\sc structure} 
\citep{Pritchardetal2000,Evannoetal2005,Wap&Gag2006,Gaoetal2011}, less has been written on how the number of 
individuals
and the number of starting allocations can affect not only the convergence of different runs of {\sc flock}, but also its 
inference of $K$. \citeauthor{Duc&Tur2012} developed estimation rules for $K$ based on simulated data 
sets with a modest number of individuals and different migration regimes and rates, numbers of loci, and $K$. 
By relying on the number of identical partitions (mean LLOD values) obtained by {\sc flock} over multiple
runs as a means of supporting 
a particular $K$,  similar---but not identical---partitions
will provide no support to a given $K$. We have shown that even at intermediate
values of $F_{ST}$ {\sc flockture} has a tendency to converge to 
different solutions where no plateaus are encountered. Despite the lack of
plateaus plots of the normalized likelihood values demonstrate that
{\sc flockture} still produces valuable information. One possible patch for this problem would be running
multiple small batches and using the programs {\sc clumpp} \citep{Jak&Ros2007} and
{\sc distruct} \citep{Rosenberg2004} to visually inspect the `best run' results, and then 
determine plateaus amongst {\sc flock}'s replicate solutions not by identical LLOD, but by visual inspection of 
partitions that
are very similar but are perhaps not identical; however such an approach would be labor-intensive
and would rely on arbitrary, difficult-to-reproduce decisions.  It is probably better to use {\sc structure} 
in such cases, keeping in mind, of course, that
estimating $K$ is a hard problem and should only be undertaken with caution, regardless of the method 
used. This is especially true in real populations which do not conform to the assumptions of genetic clustering models.

Often, newly-developed algorithms come to be understood within more general contexts and 
frameworks, and this, in turn, can help guide improvement of those algorithms.  For example, in population
genetics, the ``Markov recursion'' algorithm of \citet{Gri&Tav1994-AI} was identified by \citet{Felsensteinetal1999} 
as a special case of importance sampling \citep{Ham&Han1964}, a perspective which allowed 
\citet{Ste&Don2000} to improve upon the algorithm dramatically.  Here we have identified the {\sc flock}
algorithm as a limiting case of simulating annealing upon {\sc structure}'s no-admixture model with uncorrelated
allele frequencies. We hope this will
help users to understand and interpret {\sc flock}'s behavior and will provide insights that may be useful 
in the potential development of {\sc flock} or other clustering algorithms.

\section*{Acknowledgments}
We are grateful to Kristen Ruegg and Robin Waples for helpful comments on a draft of this paper.
P. Barry was supported by a fellowship from the Rasmuson Foundation.