### 2015-03-06 DistructPlot_tex/Fig_AllSim.tex
Added new figures, edited each section, made all figures in color and greyscale

\documentclass[varwidth=true, border={80, 10}]{standalone}
\usepackage{graphicx} 
\usepackage{amssymb} 
\usepackage{epstopdf} 
\usepackage{amsfonts} 
\usepackage{natbib} 
\usepackage{subfigure} 
\usepackage{pdfsync} 
\usepackage{xspace}

\textwidth = 6.5 in 
\textheight = 8 in 
\oddsidemargin = -0.90 in 
\evensidemargin = 0.0 in 
\topmargin = 0.0 in 
\headheight = 0.0 in 
\headsep = 0.0 in 
\parskip = 0.2in 
\parindent = 0.0in

\newlength{\nameraise} 
\setlength{\nameraise}{.07in}

\newlength{\nameoverhang} 
\setlength{\nameoverhang}{0.0in}

\begin{document} 
 A. $F_{ST}$ = \input{../SimDat17/AvgFst.txt}\\
\includegraphics{../SimDat17/final_pdf/BB_ds_Clumped_NoLabel_k005r001.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE A.C (9/9)} \\
\includegraphics{../SimDat17/final_pdf/BB_ds_Clumped_NoLabel_k005r011.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE NA.C (8/9)} \\
\includegraphics{../SimDat17/final_pdf/BB_ds_Clumped_NoLabel_k005r010.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE NA.C  (1/9)} \\
\includegraphics{../SimDat17/final_pdf/BB_ds_Clumped_NoLabel_k005r020.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE NA.NC  (7/9)} \\
\includegraphics{../SimDat17/final_pdf/BB_ds_Clumped_NoLabel_k005r022.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE NA.NC  (1/9)} \\
\includegraphics{../SimDat17/final_pdf/BB_ds_Clumped_NoLabel_k005r027.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE NA.NC  (1/9)} \\
\includegraphics{../SimDat17/final_pdf/BB_ds_Clumped_NoLabel_k005r028.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{FLOCK (9/9)} \\
\noindent\rule[.5cm]{8.2in}{0.4pt}\\
\vspace{-2.5em}\\
B. $F_{ST}$ = \input{../SimDat38/AvgFst.txt}\\
\includegraphics{../SimDat38/final_pdf/BB_ds_Clumped_NoLabel_k005r001.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE A.C (9/9)} \\
\includegraphics{../SimDat38/final_pdf/BB_ds_Clumped_NoLabel_k005r010.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE NA.C (9/9)} \\
\includegraphics{../SimDat38/final_pdf/BB_ds_Clumped_NoLabel_k005r019.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE NA.NC  (9/9)} \\
\includegraphics{../SimDat38/final_pdf/BB_ds_Clumped_NoLabel_k005r028.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{FLOCK (9/9)} \\
\noindent\rule[.5cm]{8.2in}{0.4pt}\\
\vspace{-2.5em}\\
C. $F_{ST}$ = \input{../SimDat10/AvgFst.txt}\\
\includegraphics{../SimDat10/final_pdf/BB_ds_Clumped_NoLabel_k005r001.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE A.C (9/9)} \\
\includegraphics{../SimDat10/final_pdf/BB_ds_Clumped_NoLabel_k005r010.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE NA.C (9/9)} \\
\includegraphics{../SimDat10/final_pdf/BB_ds_Clumped_NoLabel_k005r019.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE NA.NC  (9/9)} \\
\includegraphics{../SimDat10/final_pdf/BB_ds_Clumped_NoLabel_k005r028.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{FLOCK (9/9)} \\
\noindent\rule[.5cm]{8.2in}{0.4pt}\\
\vspace{-2.5em}\\
D. $F_{ST}$ = \input{../SimDat13/AvgFst.txt}\\
\includegraphics{../SimDat13/final_pdf/BB_ds_Clumped_NoLabel_k005r001.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE A.C (9/9)} \\
\includegraphics{../SimDat13/final_pdf/BB_ds_Clumped_NoLabel_k005r010.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE NA.C (9/9)} \\
\includegraphics{../SimDat13/final_pdf/BB_ds_Clumped_NoLabel_k005r019.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE NA.NC  (9/9)} \\
\includegraphics{../SimDat13/final_pdf/BB_ds_Clumped_NoLabel_k005r029.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{FLOCK (9/9)} \\
\includegraphics{../SimDat13/final_pdf/BB_ds_Clumped_NoLabel_k005r028.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{FLOCK (1/9)} \\
\includegraphics{../SimDat13/final_pdf/BB_ds_Clumped_NoLabel_k005r031.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{FLOCK (1/9)} \\
\includegraphics{../SimDat13/final_pdf/BB_ds_Clumped_NoLabel_k005r032.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{FLOCK (1/9)} \\
\noindent\rule[.5cm]{8.2in}{0.4pt}\\
\vspace{-2.5em}\\
E. $F_{ST}$ = \input{../SimDat31/AvgFst.txt}\\
\includegraphics{../SimDat31/final_pdf/BB_ds_Clumped_NoLabel_k005r001.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE A.C (9/9)} \\
\includegraphics{../SimDat31/final_pdf/BB_ds_Clumped_NoLabel_k005r010.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE NA.C (9/9)} \\
\includegraphics{../SimDat31/final_pdf/BB_ds_Clumped_NoLabel_k005r019.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{STRUCTURE NA.NC  (9/9)} \\
\includegraphics{../SimDat31/final_pdf/BB_ds_Clumped_NoLabel_k005r028.pdf}~\hspace*{-\nameoverhang}\raisebox{\nameraise}{FLOCK (1/9)*} \\

\end{document}