### 2015-03-06 tex/abstract.tex

%!TEX root = flock-comment-main.tex

      We show that the the algorithm in the program {\sc flock} \citep{Duc&Tur2009} can be
interpreted as an estimation procedure using  
a model that is essentially identical to the {\sc structure} 
\citep{Pritchardetal2000} model with no admixture and non-correlated 
allele frequency priors. The computational routine in {\sc flock} is 
equivalent to searching for the maximum-{\em a-posteriori}
estimate of this restricted {\sc structure} model via a simulated 
annealing algorithm with an extreme cooling 
schedule (namely, the exponent on the objective function~$\rightarrow \infty$).  We 
demonstrate the similarities between the two programs in a two step approach. First,
we modified the source code of  {\sc structure} to resemble {\sc flock} producing
the program {\sc flockture}. With simulated data we confirm similar behavior between 
{\sc flock}  and {\sc flockture} by assing the programs' performance
to assign individuals back to their correct population with a 
0-1 loss function (ie. a loss of 1 is incurred if an individual is assinged 
to an incorrect group). Second, we simulated many datasets under varying 
levels of population differentiation for both microsatllite and SNP genoyptes. We show that
{\sc flockture} yields results similar to {\sc structure} although with much larger 
variability from run to run. {\sc flockture} did perform better than {\sc structure} 
when genotypes were composed of SNPs and differentiation was moderate 
($F_{ST}=0.022-0.032$). When differentiation was low, {\sc structure} had a lower loss
for both marker types. Interpreting {\sc flock}'s algorithm as a special case of the model in 
{\sc structure} should aid in understanding the program's output and behavior.